Equalizer Programming Guide

### 2008-10-16 Developer/ProgrammingGuide/paper.tex
Programming Guide: update for 2D pixel compounds
[ ] May break build
[ ] Breaks existing applications
[ ] Bugfix
[ ] New Feature
[ ] Cleanup
[ ] Optimization
[x] Documentation

\date{
  \vspace{2cm}
  \begin{figure}[ht]
    \hspace{-4.5cm}\includegraphics[width=22cm]{images/teaser.pdf}
  \end{figure}
  \vspace{1cm}
%  \textbf{RELEASE CANDIDATE}\\[\medskipamount]
  Version 1.3.11 for Equalizer 0.5.5+svn\\[\medskipamount]
  \today
}

@@

\textcopyright 2007-2008 Eyescale Software GmbH. All rights reserved. No
permission is granted to copy, distribute, or create derivative works
from the contents of this electronic documentation in any manner, in
whole or in part, without the prior written permission of Eyescale
Software GmbH.

@@

If you have comments about the content, accuracy or comprehensibility of
this programming guide, please contact
\htmladdnormallink{eile@equalizergraphics.com}
{mailto:eile@equalizergraphics.com?subject=Equalizer\%20Programming\%20Guide}.

@@

\clearpage
\thispagestyle{empty}
\tableofcontents
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Version for Equalizer 0.5\\
    1.3.1   & May 02, 2008 & Add \sref{sStatistics}, improved images based on 1.2 print\\
    1.3.2   & May 05, 2008 & Add network distribution on eqPly model data\\
    1.3.3   & May 10-19, 2008 & add \aref{aFileFormat}\\
    1.3.4   & July 3, 2008 & Update \sref{sStatistics}\\
    1.3.5   & July 7, 2008 & Update after namespace naming cleanup\\
    1.3.6   & July 20, 2008 & Add accum and samples to \aref{aFileFormat}, add \sref{sDrawableConfig}\\
    1.3.7   & Aug 22, 2008 & Update \sref{sEqplyWIndow} with OSWindow changes\\
    1.3.8   & Aug 25, 2008 & Update \sref{sEventHandling} with OSWindow changes\\
    1.3.9   & Aug 29, 2008 & Update \sref{sEventHandling}: event handling cleanup\\
    1.3.10  & Oct 7, 2008  & Updated to reflect 2D pixel compounds\\
    1.3.11  & Oct 13, 2008 & Updated to reflect eq::net::Object cleanup\\
    \hline
  \end{tabularx}}
\thispagestyle{empty}
\clearpage

@@

\subsubsection{Frustum}
Compounds have a frustum description to define the physical layout of
the display environment.

@@

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/directSend.pdf}
  {\caption{\small\label{fDirectSend}Parallel Direct Send Compositing}}
\end{wrapfigure}
In direct send compositing, each rendering channel is also responsible
for the sort-last composition of one screen-space tile. He receives
the framebuffer pixels for his tile from all the other channels. The
size of one tile decreases linearly with the number of source channels,
which keeps the total amount of pixel data per channel constant.

After performing the sort-last compositing, the color information is
transferred to the destination channel, similarly to a 2D (sort-first)
compound. The amount of pixel data for this part of the compositing
pipeline also approaches a constant value, i.e., the full frame buffer.

@@

The offset of input and output frames characterizes the position of the
frame data with respect to the framebuffer, that is, the
\textbf{window's} lower-left corner. For output frames this is the
position of the channel with respect to the window.

@@

When using 2D (sort-first) or stereo decompositions, no special
programming is needed to achieve good scalability, as \textsf{eVolve} is
mostly fill-limited and therefore scales nicely in these modes.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/slices.pdf}
  {\caption{\small\label{fSlices}Blending Slices in 3D-Texture-based
      Volume Rendering}}
\end{wrapfigure}
DB compounds have the advantage of scaling any part of the volume
rendering pipeline: texture and main memory (smaller bricks for each
channel), fill rate (less samples per channel) and IO bandwidth for
time-dependent data (less data per time step and channel). Since the
amount of texture memory needed for each node decreases linearly, they
make it possible to render data sets which are not feasible to
visualize with any other approach.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \vspace{-2ex}\includegraphics[width=.382\textwidth]{images/volResult.png}
  {\caption{\small\label{fVolResult}Result of \fig{fBlendPerspective}}}
  \vspace{-2ex}
\end{wrapfigure}
For orthographic rendering, determining the compositing order of the
input frames is trivial. The screen-space orientation of the volume
bricks determines the order in which they have to be composited. The
bricks in \textsf{eVolve} are created by slicing the volume along one
dimension. Therefore the range of the resulting frame buffer images,
together with the sorting order, is used to arrange the frames during
compositing. \fig{fBlendOrtho} shows this composition for one view.

Finding the correct assembly order for perspective frustra is more
complex. The perspective distortion invalidates a simple orientation
criteria like the one used for orthographic frustra. For the view and
frustum setup shown in \fig{fBlendPerspective}\footnote{Volume Data Set
  courtesy of: AVS, USA} the correct compositing order is 4-3-1-2 or
1-4-3-2.

>>>>

\date{
  \vspace{2cm}
  \begin{figure}[ht]
    \hspace{-4.5cm}\includegraphics[width=22cm]{images/teaser.pdf}
  \end{figure}
  \vspace{1cm}
%  \textbf{RELEASE CANDIDATE}\\[\medskipamount]
  Version 1.3.12 for Equalizer 0.5.5+svn\\[\medskipamount]
  \today
}

@@

\textcopyright 2007-2008 \htmladdnormallink{Eyescale Software GmbH}
{http://www.eyescale.ch}. All rights reserved. No permission is
granted to copy, distribute, or create derivative works from the
contents of this electronic documentation in any manner, in whole or in
part, without the prior written permission of Eyescale Software GmbH.

@@

If you have comments about the content, accuracy or comprehensibility of
this programming guide, please contact
\htmladdnormallink{eile@eyescale.ch}
{mailto:eile@eyescale.ch?subject=Equalizer\%20Programming\%20Guide}.

@@

\clearpage
\thispagestyle{empty}
\tableofcontents
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Version for Equalizer 0.5\\
    1.3.1   & May 02, 2008 & Add \sref{sStatistics}, improved images based on 1.2 print\\
    1.3.2   & May 05, 2008 & Add network distribution on eqPly model data\\
    1.3.3   & May 10-19, 2008 & add \aref{aFileFormat}\\
    1.3.4   & July 3, 2008 & Update \sref{sStatistics}\\
    1.3.5   & July 7, 2008 & Update after namespace naming cleanup\\
    1.3.6   & July 20, 2008 & Add accum and samples to \aref{aFileFormat}, add \sref{sDrawableConfig}\\
    1.3.7   & Aug 22, 2008 & Update \sref{sEqplyWIndow} with OSWindow changes\\
    1.3.8   & Aug 25, 2008 & Update \sref{sEventHandling} with OSWindow changes\\
    1.3.9   & Aug 29, 2008 & Update \sref{sEventHandling}: event handling cleanup\\
    1.3.10  & Oct 7, 2008  & Updated to reflect 2D pixel compounds\\
    1.3.11  & Oct 13, 2008 & Updated to reflect eq::net::Object cleanup\\
    1.3.12  & Oct 16, 2008 & Added \sref{sView}\\
    \hline
  \end{tabularx}}
\thispagestyle{empty}
\clearpage

@@

\subsubsection{\label{sFrustum}Frustum}
Compounds have a frustum description to define the physical layout of
the display environment.

@@

\subsection{\label{sView}View Management}

A view describes the projection of the scene onto a coherent 2D area,
typically a display or window rendered by a destination channel. The
application has read and write access to all the views of a
configuration.

Changing view parameters is typically done for non-fullscreen
application windows, where the rendering is not meant to be viewed in
real-world size in an immersive environment. A typical use case is
changing the field-of-view of the rendering.

A configuration has one or more views. A view is determined by the
frustum (wall or projection description) and eye separation. A view is
uniquely assigned to one compound and possibly one destination
channel. For a description of wall and projection parameters please
refer to \sref{sFrustum}.

The application node can query all views of the \textsf{eq::Config}. It
can change all parameters of each view (wall, projection and eye
separation). The new view parameters are applied for the next frame.

Destination channels on the render clients have access to a read-only
version of their view. The view data is automatically kept up-to-date
with the application's instance by Equalizer.

The view API is not finalized and may change in the future.

\subsubsection{View Update}

The default Equalizer event handling is using the view API to maintain
the aspect ratio of destination channels after a window resize. Without
updating the wall or projection description, the rendering would become distorted.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/viewUpdate.pdf}
  {\caption{\small\label{fViewUpdate}Event Flow during a View Update}}
\end{wrapfigure}
When a window is resized, a \textsf{CHANNEL\_RESIZE} event is eventually
generated. If the corresponding channel has a view,
\textsf{Channel::processEvent} sends a \textsf{VIEW\_RESIZE} event to
the application. This event contains the identifier of the view.

The config event is dispatched to \textsf{Config::handleView\-Resize},
which uses the view identifier to find the corresponding view. Based on
the original size and wall or projection description of the view, a new
wall or projection is computed, keeping the aspect ratio and  the height
of the frustum constant. This new frustum is automatically applied by
Equalizer at the next config frame.

\fig{fViewUpdate} shows a sequence diagram of such a view update.

@@

In direct send compositing, each rendering channel is also responsible
for the sort-last composition of one screen-space tile. He receives
the framebuffer pixels for his tile from all the other channels. The
size of one tile decreases linearly with the number of source channels,
which keeps the total amount of pixel data per channel constant.

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-1ex}\includegraphics[width=.618\textwidth]{images/directSend.pdf}
  {\caption{\small\label{fDirectSend}Parallel Direct Send Compositing}}
  \vspace{-2ex}
\end{wrapfigure}
After performing the sort-last compositing, the color information is
transferred to the destination channel, similarly to a 2D (sort-first)
compound. The amount of pixel data for this part of the compositing
pipeline also approaches a constant value, i.e., the full frame buffer.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/assembly.pdf}
  {\caption{\small\label{fAssembly}Hierarchy of assembly classes}}
\end{wrapfigure}
The offset of input and output frames characterizes the position of the
frame data with respect to the framebuffer, that is, the
\textbf{window's} lower-left corner. For output frames this is the
position of the channel with respect to the window.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/slices.pdf}
  {\caption{\small\label{fSlices}Blending Slices in 3D-Texture-based
      Volume Rendering}}
\end{wrapfigure}
When using 2D (sort-first) or stereo decompositions, no special
programming is needed to achieve good scalability, as \textsf{eVolve} is
mostly fill-limited and therefore scales nicely in these modes.

@@

DB compounds have the advantage of scaling any part of the volume
rendering pipeline: texture and main memory (smaller bricks for each
channel), fill rate (less samples per channel) and IO bandwidth for
time-dependent data (less data per time step and channel). Since the
amount of texture memory needed for each node decreases linearly, they
make it possible to render data sets which are not feasible to
visualize with any other approach.

@@

For orthographic rendering, determining the compositing order of the
input frames is trivial. The screen-space orientation of the volume
bricks determines the order in which they have to be composited. The
bricks in \textsf{eVolve} are created by slicing the volume along one
dimension. Therefore the range of the resulting frame buffer images,
together with the sorting order, is used to arrange the frames during
compositing. \fig{fBlendOrtho} shows this composition for one view.

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/volResult.png}
  {\caption{\small\label{fVolResult}Result of \fig{fBlendPerspective}}}
\end{wrapfigure}
Finding the correct assembly order for perspective frustra is more
complex. The perspective distortion invalidates a simple orientation
criteria like the one used for orthographic frustra. For the view and
frustum setup shown in \fig{fBlendPerspective}\footnote{Volume Data Set
  courtesy of: AVS, USA} the correct compositing order is 4-3-1-2 or
1-4-3-2.
